---
title: "Testing SMDW Random Forest Models"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, root.dir = 
"~/GitHub/MICS_SMDW")

library(data.table)
library(tidyverse)
library(dplyr)
```
Loading (1) Multiple Indicator Cluster Survey (MICS) data for our test set as well as (2) a list with region names from the MICS surveys (HH7_region) in the test set matched with corresponding GADM areas at level 1.

```{r}
#(1)
df.MICS_test <- readHouseHoldSurveyDataForTestSet()

#(2)
matchedRegionNames_test <- readMatchedRegionNamesForTestSet()

```

We had 3 different cases when matching HH7_region names from the MICS with a GADM area level 1. We have categorized these cases in the column "Matched" as the following:
1= HH7_region is the same as GADM area level 1
2= HH7_region is the equivalent of several GADM level 1 areas together
3= several HH7_regions are the equivalent of one GADM level 1 area
14= HH7_region is similar to a GADM level 1 area but differences exist (cases:Nepal (n=3), Samoa (n=1))
24= HH7_region is similar to several GADM level 1 areas together but differences exist (Nepal n=2)

```{r}
source("~/GitHub/MICS_SMDW/Data_cleaning/MICS_Preprocessing_Dataframes/Compiling_MICS_test_and_training_sets/FunctionsForExtractingVariablesFromMICS.R", encoding = 'UTF-8')

#For case 3 we replace the HH7_region Name With Corresponding GADM level 1 Name 
df.MICS_test <- replaceHH7_regionNameWithCorrespondingGADMNAme(df.MICS_test)

matchedRegionNames_test <- replaceHH7NameWithCorrespondingGADMNAme(matchedRegionNames_test)

```

Creating indicators for Safely Managed Drinking Water (SMDW) and calculating coverage rates at HH7_region level and country level.

```{r}
source("~/GitHub/MICS_SMDW/Data_cleaning/MICS_Preprocessing_Dataframes/CreatingSMDWIndicator.R")

df.MICS_SMDW_FullTestSet <- creatingSMDWIndicators_forTestSet(df.MICS_test)

SMDW_test_HH7 <- createIndicatorForRegionalSMDWCoverage_forTestSet(df.MICS_SMDW_FullTestSet)

SMDW_test_country <- createIndicatorForCountrySMDWCoverage(df.MICS_SMDW_FullTestSet)

```
Next we match our SMDW predictions at GADM area level 1 with our calculated SMDW coverage rates at a) HH7_region level and b) country level. We do this for predictions generated from our best performing random forest model built on the original training set as well as for the average prediction of 100 random forest models build on geographically stratified training sets.

```{r}

#Load predicted values from highest performing model and average of geographically stratified models

SMDW_Predictions <- read.csv( "~/GitHub/MICS_SMDW/Prediction/SMDWPredictions_forTest.csv", fileEncoding = "UTF-8") %>% 
  select(predict.x, mean,GID_1, GID_0)

#Load and join population counts for 2020 to prediction data frame
Population_2020 <- read.csv(here("Data/EnvironmentalFeatures/FeaturesAtGADMlevel1.csv"), encoding = "UTF-8") %>% 
  select(GID_1, GPWv4_Population_2020)

SMDW_Predictions <- SMDW_Predictions %>% 
  left_join(Population_2020, by = c("GID_1"="GID_1"))

#Join prediction data frame with matched HH7_region names

SMDW_Predictions <- SMDW_Predictions %>% 
  left_join(matchedRegionNames_test, by = c("GID_1"="GID_1")) %>% na.omit(SMDW_Predictions$HH7)

#Calculate the predicted proportion of people using SMDW in HH7_regions using predictions at GADM 1 and population estimates. 

#1) predicted values * population at GADM level 1

SMDW_Predictions$numberPeopleWithSMDW_GADM1 <- 
  SMDW_Predictions$predict.x*SMDW_Predictions$GPWv4_Population_2020

SMDW_Predictions$averageNumberPeopleWithSMDW_GADM1 <- 
  SMDW_Predictions$mean*SMDW_Predictions$GPWv4_Population_2020


#2a) calculate number of people at HH7_region using SMDW as well as the total number of people in at HH7_region
SMDW_originalPrediction_HH7 <- SMDW_Predictions %>% 
  group_by(HH7) %>% 
  mutate(numberPeopleWithSMDW_HH7 = sum(numberPeopleWithSMDW_GADM1),
         GPWv4_Population_2020_HH7 = sum(GPWv4_Population_2020))

SMDW_100Predictions_HH7 <- SMDW_Predictions %>% 
  group_by(HH7) %>% 
  mutate(numberPeopleWithSMDW_HH7 = sum(averageNumberPeopleWithSMDW_GADM1),
         GPWv4_Population_2020_HH7 = sum(GPWv4_Population_2020))


#2b)
SMDW_originalPrediction_country <- SMDW_Predictions %>% 
  group_by(country) %>% 
  mutate(numberPeopleWithSMDW_country = sum(numberPeopleWithSMDW_GADM1),
         GPWv4_Population_2020_country = sum(GPWv4_Population_2020))

SMDW_100Predictions_country <- SMDW_Predictions %>% 
  group_by(country) %>% 
  mutate(numberPeopleWithSMDW_country = sum(averageNumberPeopleWithSMDW_GADM1),
         GPWv4_Population_2020_country = sum(GPWv4_Population_2020))


#3a) calculate proportion of people using SMDW in each HH7_region

SMDW_originalPrediction_HH7$predict_HH7 <-
  SMDW_originalPrediction_HH7$numberPeopleWithSMDW_HH7/SMDW_originalPrediction_HH7$GPWv4_Population_2020_HH7

SMDW_100Predictions_HH7$predict_HH7 <-
  SMDW_100Predictions_HH7$numberPeopleWithSMDW_HH7/SMDW_100Predictions_HH7$GPWv4_Population_2020_HH7


#3a) calculate proportion of people using SMDW in each country

SMDW_originalPrediction_country$predict_country <-
  SMDW_originalPrediction_country$numberPeopleWithSMDW_country/SMDW_originalPrediction_country$GPWv4_Population_2020_country

SMDW_100Predictions_country$predict_country <-
  SMDW_100Predictions_country$numberPeopleWithSMDW_country/SMDW_100Predictions_country$GPWv4_Population_2020_country


#Join SMDW predictions with test set at HH7_region and country level

SMDW_test_HH7_original <- SMDW_originalPrediction_HH7 %>% 
  left_join(SMDW_test_HH7, by = c("HH7"="HH7_region"),all=TRUE) %>% 
  select("predict_HH7","SMDWcoverageAtRegionalLevel","HH7","country", "HouseholdsInRegion.Freq.x", "GID_0") %>% 
  unique() %>%
  na.omit()

SMDW_test_HH7_100pred <- SMDW_100Predictions_HH7 %>% 
  left_join(SMDW_test_HH7, by = c("HH7"="HH7_region"),all=TRUE) %>% 
  select("predict_HH7","SMDWcoverageAtRegionalLevel","HH7","country", "HouseholdsInRegion.Freq.x", "GID_0") %>% 
  unique() %>%
  na.omit()


SMDW_test_country_original <- SMDW_originalPrediction_country %>% 
  left_join(SMDW_test_country, by = c("country"="country"),all=TRUE) %>% 
  select("predict_country","SMDWcoverageAtCountryLevel","country", "HouseholdsInCountry.Freq.x", "GID_0") %>% 
  unique() %>%
  na.omit()

SMDW_test_country_100pred <- SMDW_100Predictions_country %>% 
  left_join(SMDW_test_country, by = c("country"="country"),all=TRUE) %>% 
  select("predict_country","SMDWcoverageAtCountryLevel","country", "HouseholdsInCountry.Freq.x", "GID_0") %>% 
  unique() %>%
  na.omit()

# calculate coefficient of determination and mean square error

#for single best performing model
r2_test_HH7_original <- 1 - (sum((SMDW_test_HH7_original$SMDWcoverageAtRegionalLevel-SMDW_test_HH7_original$predict_HH7)^2)/sum((SMDW_test_HH7_original$SMDWcoverageAtRegionalLevel-mean(SMDW_test_HH7_original$SMDWcoverageAtRegionalLevel))^2))
r2_test_HH7_original

#r2= 1 - (sum((actual-predicted)^2)/sum((actual-mean(actual))^2))

MAE_test_HH7_original <- sum(abs(SMDW_test_HH7_original$SMDWcoverageAtRegionalLevel - SMDW_test_HH7_original$predict_HH7))/length(SMDW_test_HH7_original$SMDWcoverageAtRegionalLevel)
MAE_test_HH7_original

#for 100 geographically stratified models
r2_test_HH7_100pred <- 1 - (sum((SMDW_test_HH7_100pred$SMDWcoverageAtRegionalLevel-SMDW_test_HH7_100pred$predict_HH7)^2)/sum((SMDW_test_HH7_100pred$SMDWcoverageAtRegionalLevel-mean(SMDW_test_HH7_100pred$SMDWcoverageAtRegionalLevel))^2))
r2_test_HH7_100pred

MAE_test_HH7_100pred <- sum(abs(SMDW_test_HH7_100pred$SMDWcoverageAtRegionalLevel - SMDW_test_HH7_100pred$predict_HH7))/length(SMDW_test_HH7_100pred$SMDWcoverageAtRegionalLevel)
MAE_test_HH7_100pred

r2_test_country_original <- 1 - (sum((SMDW_test_country_original$SMDWcoverageAtCountryLevel-SMDW_test_country_original$predict_country)^2)/sum((SMDW_test_country_original$SMDWcoverageAtCountryLevel-mean(SMDW_test_country_original$SMDWcoverageAtCountryLevel))^2))
r2_test_country_original

MAE_test_country_original <- sum(abs(SMDW_test_country_original$SMDWcoverageAtCountryLevel - SMDW_test_country_original$predict_country))/length(SMDW_test_country_original$SMDWcoverageAtCountryLevel)
MAE_test_country_original

r2_test_country_100pred <- 1 - (sum((SMDW_test_country_100pred$SMDWcoverageAtCountryLevel-SMDW_test_country_100pred$predict_country)^2)/sum((SMDW_test_country_100pred$SMDWcoverageAtCountryLevel-mean(SMDW_test_country_100pred$SMDWcoverageAtCountryLevel))^2))
r2_test_country_100pred

MAE_test_country_100pred<- sum(abs(SMDW_test_country_100pred$SMDWcoverageAtCountryLevel - SMDW_test_country_100pred$predict_country))/length(SMDW_test_country_100pred$SMDWcoverageAtCountryLevel)
MAE_test_country_100pred
```

Plotting test set at HH7

```{r}
# Load data frame with country Income Group

source("~/GitHub/MICS_SMDW/Data_cleaning/MICS_Preprocessing_Dataframes/Compiling_MICS_test_and_training_sets/functionsForJoiningAndStructuringDataframes.R")

country_Income_Groups <- readCountryIncomeGroup() %>% 
  select("Code", "Income.group")

GlobalRegions <- readAndRenameStandardGlobalRegionNames()
  
SMDW_test_HH7_original <- SMDW_test_HH7_original %>%
  left_join(country_Income_Groups, by = c("GID_0"="Code")) %>%
  left_join(GlobalRegions, by = c("GID_0"="ISO.alpha3.Code"))

SMDW_test_HH7_100pred <- SMDW_test_HH7_100pred %>%
  left_join(country_Income_Groups, by = c("GID_0"="Code")) %>%
  left_join(GlobalRegions, by = c("GID_0"="ISO.alpha3.Code"))

#SMDW_test_HH7_original$Income.group <-as.factor(SMDW_test_HH7_original$Income.group)

#plotting distibution of income level and UN regions in test set

table(SMDW_test_HH7_original$Income.group)

Plot_IncomeLevelDistribution_Test <- ggplot(SMDW_test_HH7_original,   aes(x=as.factor(Income.group), y=..count.. / sum(..count..))) +
  ggtitle("Income Group Distribution in Test Set")+
  geom_bar(fill="light blue") + 
  labs(x = "",
       y = "Proportion of subnational regions")+
  theme_bw()
Plot_IncomeLevelDistribution_Test

table(SMDW_test_HH7_original$UNRegionName)

Plot_UNRegionDistribution_Test <- ggplot(SMDW_test_HH7_original, aes(x=as.factor(UNRegionName), y=..count.. / sum(..count..))) +
  ggtitle("UN Region Distribution in Test Set")+
  geom_bar(fill="light blue") + 
  labs(x = "",
       y = "Proportion of subnational regions")+
  theme_bw()
Plot_UNRegionDistribution_Test

```

```{r}
# Plot predicted vs observed values- coloured by UN Region

colours <- c("#E64B35FF", "#91D1C2FF","#3C5488FF","#00A087FF", "#4DBBD5FF")                                    
plotPredObs_SMDWtest_HH7_orig <- ggplot(SMDW_test_HH7_original, 
                      aes(x = predict_HH7, 
                          y = SMDWcoverageAtRegionalLevel, size = HouseholdsInRegion.Freq.x, colour= UNRegionName)) +
  geom_point(shape=1) +
  labs(x = "Predicted Proportion of Population Using SMDWs (single model)",
       y = "Proportion of Population Using SMDWs") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  stat_smooth(se = T,
              colour = "black",
              size = 0.5,
              method = "lm") +
  labs(size="Households in region (n)", 
       col="UN Region")+
  scale_colour_manual(values = colours)+
  theme_bw()

plotPredObs_SMDWtest_HH7_orig


plotPredObs_SMDWtest_HH7_100pred <- ggplot(SMDW_test_HH7_100pred, 
                      aes(x = predict_HH7, 
                          y = SMDWcoverageAtRegionalLevel, size = HouseholdsInRegion.Freq.x, colour= UNRegionName)) +
  geom_point(shape=1) +
  labs(x = "Mean Predicted Proportion of Population Using SMDWs (100 models)",
       y = "Proportion of Population Using SMDWs") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  stat_smooth(se = T,
              colour = "black",
              size = 0.5,
              method = "lm") +
  labs(size="Households in region (n)", 
       col="UN Region")+
  scale_colour_manual(values = colours)+
  theme_bw()

plotPredObs_SMDWtest_HH7_100pred

```

```{r}
# Plot predicted vs observed values- coloured by income group
                                    
plotPredObs_SMDWtest_HH7_orig_ <- ggplot(SMDW_test_HH7_original, 
                      aes(x = predict_HH7, 
                          y = SMDWcoverageAtRegionalLevel, size = HouseholdsInRegion.Freq.x, colour= Income.group)) +
  geom_point(shape=1) +
  labs(x = "Predicted proportion of subnational population using SMDWs",
       y = "Proportion of subnational population using SMDWs") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  stat_smooth(se = T,
              colour = "black",
              size = 0.5,
              method = "lm") +
  labs(size="Households in region (n)", 
       col="Income Group")+
  scale_colour_manual(values = colours)+
  theme_bw()

plotPredObs_SMDWtest_HH7_orig_

plotPredObs_SMDWtest_HH7_100pred_ <- ggplot(SMDW_test_HH7_100pred, 
                      aes(x = predict_HH7, 
                          y = SMDWcoverageAtRegionalLevel, size = HouseholdsInRegion.Freq.x, colour= Income.group)) +
  geom_point(shape=1) +
  labs(x = "Predicted proportion of subnational population using SMDWs",
       y = "Proportion of subnational population using SMDWs") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  stat_smooth(se = T,
              colour = "black",
              size = 0.5,
              method = "lm") +
  labs(size="Households in region (n)", 
       col="Income Group")+
  scale_colour_manual(values = colours)+
  theme_bw()

plotPredObs_SMDWtest_HH7_100pred_
```
```
Plotting test set at country
```{r}
# Load data frame with country Income Group

SMDW_test_country_original <- SMDW_test_country_original %>%
  left_join(country_Income_Groups, by = c("GID_0"="Code")) %>%
  left_join(GlobalRegions, by = c("GID_0"="ISO.alpha3.Code"))
  
SMDW_test_country_100pred <- SMDW_test_country_100pred %>%
  left_join(country_Income_Groups, by = c("GID_0"="Code")) %>%
  left_join(GlobalRegions, by = c("GID_0"="ISO.alpha3.Code"))


# Plot predicted vs observed values- coloured by UN Region
                                    
plotPredObs_SMDWtest_country_original <- ggplot(SMDW_test_country_original, 
                      aes(x = predict_country, 
                          y = SMDWcoverageAtCountryLevel, size = HouseholdsInCountry.Freq.x, colour= UNRegionName)) +
  geom_point(shape=1) +
  labs(x = "Predicted proportion of national population using SMDWs",
       y = "Proportion of national population using SMDWs") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  stat_smooth(se = T,
              colour = "black",
              size = 0.5,
              method = "lm") +
              labs(size="Households in region (n)", 
       col="UN Region")+
  scale_colour_manual(values = colours)+
  theme_bw()

plotPredObs_SMDWtest_country_original 


plotPredObs_SMDWtest_country_100pred <- ggplot(SMDW_test_country_100pred, 
                      aes(x = predict_country, 
                          y = SMDWcoverageAtCountryLevel, size = HouseholdsInCountry.Freq.x, colour= UNRegionName)) +
  geom_point(shape=1) +
  labs(x = "Predicted proportion of national population using SMDWs",
       y = "Proportion of national population using SMDWs") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  stat_smooth(se = T,
              colour = "black",
              size = 0.5,
              method = "lm") +
              labs(size="Households in region (n)", 
       col="UN Region")+
  scale_colour_manual(values = colours)+
  theme_bw()

plotPredObs_SMDWtest_country_100pred 
```


```{r}
# Plot predicted vs observed values- coloured by country

plotPredObs_SMDWtest_country_income <- ggplot(SMDW_test_country, 
                      aes(x = predict_country, 
                          y = SMDWcoverageAtCountryLevel, size = HouseholdsInCountry.Freq.x, colour= `Income group`)) +
  geom_point(shape=1) +
  labs(x = "Predicted",
       y = "Observed") +
   coord_cartesian(xlim = c(0,0.8), ylim = c(0,0.8)) +
  geom_abline(slope = 1,
              intercept = 0,
              na.rm = FALSE, 
              show.legend = NA, 
              linetype = "dashed") +
  theme_bw()

plotPredObs_SMDWtest_country_income

```