---
title: "cluster_analysis"
author: "Esther"
date: "07/02/2022"
output: html_document
---

```{r}
# Set working directory
knitr::opts_chunk$set(echo = TRUE, root.dir = 
"./3_Feature_selection")

library(data.table)
library(factoextra)
library(tidyverse)
library(Hmisc)

```

## Pre-selected environmental varibales

We pre-elected 117 variables based on their relevance for modelling safely managed drinking water as anthropogenic, hydrogeologic, climatic, topographic and biogeographic indicators from a list of 384 variables used in ecological modelling (compiled by the ETHZ Crowther Lab). 

```{r}
bandNames <- c(
   "Abs_Lat",
  "CGIAR_Aridity_Index",
  "CGIAR_PET",
  "CHELSA_BIO_Annual_Mean_Temperature",
  "CHELSA_BIO_Annual_Precipitation",
  "CHELSA_BIO_Isothermality",
  "CHELSA_BIO_Max_Temperature_of_Warmest_Month",
  "CHELSA_BIO_Mean_Diurnal_Range",
  "CHELSA_BIO_Mean_Temperature_of_Coldest_Quarter",
  "CHELSA_BIO_Mean_Temperature_of_Driest_Quarter",
  "CHELSA_BIO_Mean_Temperature_of_Warmest_Quarter",
  "CHELSA_BIO_Mean_Temperature_of_Wettest_Quarter",
  "CHELSA_BIO_Min_Temperature_of_Coldest_Month",
  "CHELSA_BIO_Precipitation_Seasonality",
  "CHELSA_BIO_Precipitation_of_Coldest_Quarter",
  "CHELSA_BIO_Precipitation_of_Driest_Month",
  "CHELSA_BIO_Precipitation_of_Driest_Quarter",
  "CHELSA_BIO_Precipitation_of_Warmest_Quarter",
  "CHELSA_BIO_Precipitation_of_Wettest_Month",
  "CHELSA_BIO_Precipitation_of_Wettest_Quarter",
  "CHELSA_BIO_Temperature_Annual_Range",
  "CHELSA_BIO_Temperature_Seasonality",
  "CIFOR_TropicalPeatlandDepth",
  "CIFOR_TropicalPeatlandExtent",
  "CMS_HeterotrophicRespiration_usingEquationFromBondLamberty2004",
  "CMS_HeterotrophicRespiration_usingEquationFromSubke2006",
  "CMS_SoilRespiration_mean",
  "CSP_Global_Human_Modification",
  "ConsensusLandCoverClass_Barren",
  "ConsensusLandCoverClass_Cultivated_and_Managed_Vegetation",
  "ConsensusLandCoverClass_Deciduous_Broadleaf_Trees",
  "ConsensusLandCoverClass_Evergreen_Broadleaf_Trees",
  "ConsensusLandCoverClass_Evergreen_Deciduous_Needleleaf_Trees",
  "ConsensusLandCoverClass_Herbaceous_Vegetation",
  "ConsensusLandCoverClass_Mixed_Other_Trees",
  "ConsensusLandCoverClass_Open_Water",
  "ConsensusLandCoverClass_Regularly_Flooded_Vegetation",
  "ConsensusLandCoverClass_Shrubs",
  "ConsensusLandCoverClass_Snow_Ice",
  "ConsensusLandCoverClass_Urban_Builtup",
  "ConsensusLandCover_Human_Development_Percentage",
  "CrowtherLab_Tree_Density",
  "EarthEnvCloudCover_CloudForestPrediction",
  "EarthEnvCloudCover_Mean",
  "EarthEnvTexture_CoOfVar_EVI",
  "EarthEnvTexture_Contrast_EVI",
  "EarthEnvTexture_Correlation_EVI",
  "EarthEnvTexture_Dissimilarity_EVI",
  "EarthEnvTexture_Entropy_EVI",
  "EarthEnvTexture_Evenness_EVI",
  "EarthEnvTexture_Homogeneity_EVI",
  "EarthEnvTexture_Maximum_EVI",
  "EarthEnvTexture_Range_EVI",
  "EarthEnvTexture_Shannon_Index",
  "EarthEnvTexture_Simpson_Index",
  "EarthEnvTexture_Std_EVI",
  "EarthEnvTexture_Uniformity_EVI",
  "EarthEnvTexture_Variance_EVI",
  "EarthEnvTopoMed_1stOrderPartialDerivEW",
  "EarthEnvTopoMed_1stOrderPartialDerivNS",
  "EarthEnvTopoMed_2ndOrderPartialDerivEW",
  "EarthEnvTopoMed_2ndOrderPartialDerivNS",
  "EarthEnvTopoMed_AspectCosine",
  "EarthEnvTopoMed_AspectSine",
  "EarthEnvTopoMed_Eastness",
  "EarthEnvTopoMed_Elevation",
  "EarthEnvTopoMed_Northness",
  "EarthEnvTopoMed_ProfileCurvature",
  "EarthEnvTopoMed_Roughness",
  "EarthEnvTopoMed_Slope",
  "EarthEnvTopoMed_TangentialCurvature",
  "EarthEnvTopoMed_TerrainRuggednessIndex",
  "EarthEnvTopoMed_TopoPositionIndex",
  "EarthEnvTopoMed_VectorRuggednessMeasure",
  "EsaCci_BurntAreasProbability",
  "FanEtAl_Depth_to_Water_Table_AnnualMean",
  "FanEtAl_Depth_to_Water_Table_AnnualSD",
  "GHS_Population_Density",
  "GLW3_RuminantsDistribution_downsampled10km",
  "GPWv4_Population_Density",
  "GiriEtAl_MangrovesExtent",
  "GoshEtAl_GlobalGDP2006",
  "IPCC_Global_Biomass",
  "MODIS_EVI",
  "MODIS_FPAR",
  "MODIS_GPP",
  "MODIS_LAI",
  "MODIS_NDVI",
  "MODIS_NPP",
  "MODIS_Nadir_Reflectance_Band1",
  "MODIS_Nadir_Reflectance_Band2",
  "MODIS_Nadir_Reflectance_Band3",
  "MODIS_Nadir_Reflectance_Band4",
  "MODIS_Nadir_Reflectance_Band5",
  "MODIS_Nadir_Reflectance_Band6",
  "MODIS_Nadir_Reflectance_Band7",
  "NASA_ForestCanopyHeight",
  "PelletierEtAl_SoilAndSedimentaryDepositThicknesses",
  #"Resolve_Biome",
  #"Resolve_Ecoregion",
  "SG_Absolute_depth_to_bedrock",
  "SG_Bulk_density_015cm",
  "SG_CEC_015cm",
  "SG_Clay_Content_015cm",
  "SG_Coarse_fragments_015cm",
"SG_Depth_to_bedrock",
  "SG_H2O_Capacity_015cm",
  "SG_SOC_Content_015cm",
  "SG_SOC_Density_015cm",
  "SG_SOC_Stock_005cm_to_015cm",
  "SG_Saturated_H2O_Content_015cm",
  "SG_Silt_Content_015cm",
  "SG_Soil_H2O_Capacity_pF_20_015cm",
  "SG_Soil_pH_H2O_015cm",
  "SpawnEtAl_HarmonizedAGBiomass",
  "SpawnEtAl_HarmonizedBGBiomass",
  "TootchiEtAl_WetlandsRegularlyFlooded",
  "UNEP_WCMC_Biomass",
  "WCS_Human_Footprint_2009"
  #"WWF_Biome",
  #"WWF_Ecoregion"
)

```

Load 100 000 randomly sampled data points from the selected features layers and then exclude observations with NAs. 
```{r}
# Load data from 100k randomly sampled data points
random_points_data <- fread("./switchdrive/Data_for_reviewers/EnvironmentalFeatures/subs100k.csv") %>% mutate_all(as.numeric)

#select variables 
random_points <- random_points_data %>% select(all_of(bandNames)) %>%
  na.omit() #loose about 20000 observations which have NAs
```

## Use clustering of variables to reduce the number of features for our model.
```{r}

library(ClustOfVar)
hclust <- hclustvar(random_points %>% sample_n(50000)) 
plot(hclust)

#We select 24, 39, 51, 73 and 91 clusters
for (k in c(24,39,51,73,91)){
  cuttree <- cutreevar(hclust, k = k)
  
  df <- cuttree$cluster %>% as.data.frame() %>% rename('cluster' = 1) %>% rownames_to_column("variable") %>% arrange(cluster)
  df %>% fwrite(paste0('./Feature Selection/clustofvar_50k_new_',k,'.csv'))
  
}
```

